\thispagestyle{empty}
\noindent \textbf{Abstract}\\
\newline


\\
\newline	
\noindent
In this project we have investigated and assessed wireless patient monitoring solutions. Trough qualitative interviews we have investigated the infrastructure and practice of wireless ECG monitoring systems at two Norwegian hospitals. Combining what we learned from the existing solutions with different WBAN implementations found in literature, we propose a design for a wireless monitoring system based on available technology and open standards. We validated this design by building prototype that test certain critical aspects of wireless ECG monitoring. The prototype was based on a Bluetooth Smart enabled node, an Android gateway and a modern web-server for real time streaming.

In conclusion we see that the current version of Bluetooth Smart is able to support continuous streaming of clinical grade raw ECG at a 1000 Hz sampling rate with 32 bit samples. A BLE enabled ECG device can support this throughput and still have more than 4 times the battery lifetime compared to the existing solutions we investigated. We have also assessed the end-to-end delay in our proposed design for a monitoring solution, and showed that for one of our use cases, this delay averages around 60ms.

We have evaluated the technology backing a WBAN enabled hospital as ready for clinical usage, but don't believe the current organization of clinical technology will change before outside factors make the manufacturers rethink their current business models. 