\newglossaryentry{EMR}
{
    name=EMR,
    description={Electronic Medical Record}
}

\newglossaryentry{PAS}
{
    name=PAS,
    description={Patient Administration System}
}

\newglossaryentry{RIS}
{
    name=RIS,
    description={Radiology Information System. This is the EMR the radiology department uses}
}

\newglossaryentry{HIS}
{
    name=HIS,
    description={Hospital Information Systems}
}

\newglossaryentry{DICOM}
{
    name=DICOM,
    description={A file format and network communications protocol that is used to transfer different medical information online}
}

\newglossaryentry{DIPS}
{
    name=DIPS,
    description={DIPS ASA is the largest supplier of electronic patient record systems to Norwegian hospitals}
}

\newglossaryentry{Hash}
{
    name=Cryptographic hash,
    description={A cryptographic hash is a hash function that takes an arbitrary block of data and returns a fixed-size bit string}
}

\newglossaryentry{SSN}
{
    name=SSN,
    description={Social Security Number. Number used to uniquely identify individuals}
}

\newglossaryentry{KITH}
{
    name=KITH,
    description={Norwegian Centre for Informatics in Health and Social Care}
}

\newglossaryentry{bruteforce}
{
    name=Brute force,
    description={Brute force consists of systematically checking all possible keys or passwords until the correct one is found}
}

\newglossaryentry{Radiologist}
{
    name=Radiologist,
    description={Radiologist are doctors with training and education in performing and diagnosing radiology examinations. They are often assisted by a radiography technician in performing the examinations}
}

\newglossaryentry{Radiography technician}
{
    name=Radiographer,
    description={Also known as a radiographer, is a technician that works directly with examining patients at a radiology department. They operate hardware and perform examinations, but do not diagnose the results (see radiologist)}
}

\newglossaryentry{JIRA}
{
    name=JIRA,
    description={An agile development tool. Helps organize sprints and works as a issue tracker}
}

\newglossaryentry{Confluence}
{
    name=Confluence,
    description={A Wiki used to organize meeting notes, research information, weekly reports and sprint retrospects}
}

\newglossaryentry{Scrum}
{
    name=Scrum,
    description={An agile development methodology where frequent short meetings are important and work is divided into sprints}
}

\newglossaryentry{Sprint}
{
    name=Sprint,
    description={A collective effort to finish a set of issues. In our project a sprint goes along a two-week period}
}

\newglossaryentry{Capgemeni}
{
    name=Capgemeni,
    description={The customer and main point of contact. Is acting on GE Healthcare's behalf and considered product owner}
}

\newglossaryentry{ldap}
{
    name=LDAP,
    description={Lightweight Directory Access Protocol}
}

\newglossaryentry{xml}
{
    name=XML,
    description={Extensible Markup Language. Universal markup language for sharing of structured data between information systems}
}

\newglossaryentry{aidl}
{
    name=AIDL,
    description={Android Interface Definition Language}
}

\newglossaryentry{mvc}
{
    name=MVC,
    description={Model–view–controller. A software architecture pattern commonly used when designing user interfaces. The pattern divides an application into three connected parts; model, view and controller}
}

\newglossaryentry{sharedpreferences}
{
    name=SharedPreferences,
    description={An Android specific location where one can store data in a \emph{key, value, pair} format}
}

\newglossaryentry{aes}
{
    name=AES,
    description={Advanced Encryption Standard}
}

\newglossaryentry{PACS}
{
    name=PACS,
    description={A medical picture archiving and communication system}
}

\newglossaryentry{Inflate}
{
    name=Inflate,
    description={Inflate is the process of rendering XML specific GUI elements and creating the views specified within it}
}

\newglossaryentry{GEHealthcare}
{
    name=GE Healthcare,
    description={GE Healthcare is considered a customer.}
}

\newglossaryentry{LIS}
{
    name=LIS,
    description={Laboratory Information Systems}
}

\newglossaryentry{konik}
{
    name=Konik,
    description={Norwegian consultant firm based in Oslo. Specializes in Microsoft technology and hospital integration}
}

\newglossaryentry{OUS}
{
    name=OUS,
    description={Oslo University Hospital}
}

\newglossaryentry{PDF}
{
    name=PDF,
    description={Portable Document Format. File format used to present documents in a platform independent manner}
}